Identifying Software Performance Changes Across Variants and Versions

### 2020-08-11 paper.tex
added ACM copyright

@@

	Our main idea is as follows: We address the configuration complexity of this problem by selecting representative sample sets of configurations.
	Next, we uniformly sample a constant number of commits for each configuration and conduct respective performance measurements.
	Based on these measurement data, we assemble a prediction model.
	For each configuration, we estimate the likelihood of each commit being a change point, i.e. that performance of some configurations changes abruptly compared to the previous commit.
	Next, we leverage similarities in the performance histories of configurations that share common options.
	Since we sample the commits for each configuration independently, for each change point, we obtain many estimations using measurements of different commits.
	Overall, this allows us to obtain more accurate estimations of performance-changing commits with tractable effort.
	Based on a mapping from configurations to predicted change point probabilities for each configuration and commit, we derive the options responsible for each particular change point: configuration-dependent change points. {\color{red}g}
	In summary, we offer the following contributions:
	\begin{itemize}
		\item A {\color{red}novel perspective and} technique to effectively identify shifts in performance of configurable software systems. We are able to pinpoint causative commits and affected configuration options with high accuracy and tractable effort.
		\item A feasibility demonstration of our approach by implementing an adaptive learning strategy to obtain accurate estimations with acceptable measurement cost.
		\item An evaluation using both synthetic and real-world performance data from three configurable software systems. Synthetic data lets us assess our model and approach conceptually and at scale, whereas we are able to assess practicality with real-world data.
		\item A companion website\footnote{\url{https://github.com/AI-4-SE/Changepoints-Across-Variants-And-Versions}} providing supplementary material including a reference implementation of our approach, performance measurement data, and additional visualizations.
	\end{itemize}

@@

	{\color{red}In our latter experiment, we emulated our approach on a broad but finite set of measurements rather than setting it up as an online pipeline. This does not correspond to the original layout of our algorithm, but lets us efficiently explore some limitations of this approach. We have learned, that for interpreting the results of our approach, domain knowledge is advantageous as we have identified incomplete/inconsistent configuration interfaces as possible limiting factors. We acknowledge that, while our approach performs well as an automated pipeline, for a real-world applications, a semi-automated setup with additional manual measurement prioritization and knowledge about configuration constraints would improve the applicability of our approach.} This is a promising further avenue of future work.

@@

	\section{Related Work}\label{sec:rw}
	\paragraph{Performance-relevant Commits} Identifying performance-changing commits is a prevalent challenge in regression testing, often applying change point detection algorithms on batches of performance observations~\cite{cityIdentifying2014,daly_industry_2020}. A way to reduce testing overhead is to employ only a fraction of performance observations or prioritize commits. Sandoval Alcocer et al. assess performance for uniformly selected commits to estimate the risk of a performance change introduced by unobserved commits~\cite{sandoval_alcocer_learning_2016,alcocer_prioritizing_2020}. Huang et al. estimate the risk of performance changes solely based on static analysis of individual commits to prioritize commits for regression testing without assessing performance~\cite{huang_performance_2014}. Mühlbauer et al. use Gaussian Processes (GP) and iterative sampling to learn performance histories with few performance observations~\cite{muhlbauer_accurate_2019}. They extract performance changing commits by applying change point detection algorithms to learned performance histories. All of the above approaches reduce testing overhead, but do not address the performance across different configurations of software systems. Our approach incorporates performance changes across different configurations by leveraging similarities in related configuration's performance histories.
	\paragraph{Performance of Configurable Software Systems} Associating configuration options and interactions with performance a conceptual basis for our work. There exists extensive work on modeling configuration options as features for machine learning, such as classification and regression trees~\cite{guoVariabilityawarePerformancePrediction2013,sarkarCostEfficientSamplingPerformance,nairUsingBadLearners2017}, multi-variable regression~\cite{siegmundPerformanceinfluenceModelsHighly2015}, and deep neural networks~\cite{haDeepPerf2019}. In contrast to learning performance for arbitrary configurations, our approach aims at explaining performance change, and subsequently, a change in a configuration option's or interaction's influence on performance over time. A similar scenario to ours has been studied by Jamshidi et al.~\cite{jamishidi_transfer_2017,jamshidi_transfer_gp_2017,jamshidi_learning_2018}. Between software versions, the performance influence of only few configuration options changes. The authors propose two approaches for learning a transfer function between performance models of different environments, such as versions, based on GPs~\cite{jamshidi_transfer_gp_2017} and progressive shrinking the configuration space~\cite{jamshidi_learning_2018}. We see great potential for future work in applying GPs to our problem statement as they have been successfully applied for the two orthogonal perspectives of time and configuration space.

>>>>

	\copyrightyear{2020}
	\acmYear{2020}
	\setcopyright{acmcopyright}\acmConference[ASE '20]{35th IEEE/ACM International Conference on Automated Software Engineering}{September 21--25, 2020}{Virtual Event, Australia}
	\acmBooktitle{35th IEEE/ACM International Conference on Automated Software Engineering (ASE '20), September 21--25, 2020, Virtual Event, Australia}
	\acmPrice{15.00}
	\acmDOI{10.1145/3324884.3416573}
	\acmISBN{978-1-4503-6768-4/20/09}

@@

	{\color{red}We propose a novel approach, \emph{spatio-temporal performance analysis}}. Our main idea is as follows: We address the configuration complexity of this problem by selecting representative sample sets of configurations.
	Next, we uniformly sample a constant number of commits for each configuration and conduct respective performance measurements.
	Based on these measurement data, we assemble a prediction model.
	For each configuration, we estimate the likelihood of each commit being a change point, i.e. that performance of some configurations changes abruptly compared to the previous commit.
	Next, we leverage similarities in the performance histories of configurations that share common options.
	Since we sample the commits for each configuration independently, for each change point, we obtain many estimations using measurements of different commits.
	Overall, this allows us to obtain more accurate estimations of performance-changing commits with tractable effort.
	Based on a mapping from configurations to predicted change point probabilities for each configuration and commit, we derive the options responsible for each particular change point: configuration-dependent change points.
	In summary, we offer the following contributions:
	\begin{itemize}
		\item A {\color{red}novel perspective and} technique to effectively identify shifts in performance of configurable software systems. We are able to pinpoint causative commits and affected configuration options with high accuracy and tractable effort.
		\item A feasibility demonstration of our approach by implementing an adaptive learning strategy to obtain accurate estimations with acceptable measurement cost.
		\item An evaluation using both synthetic and real-world performance data from three configurable software systems. Synthetic data lets us assess our model and approach conceptually and at scale, whereas we are able to assess practicality with real-world data.
		\item A companion website\footnote{\url{https://github.com/AI-4-SE/Changepoints-Across-Variants-And-Versions} or archived at {\color{red}\url{SOFTWAR-HERITAGE}}} providing supplementary material including a reference implementation of our approach, performance measurement data, and additional visualizations.
	\end{itemize}

@@

	{\color{red}In our latter experiment, we emulated our approach on a broad but finite set of measurements rather than setting it up as an online pipeline. This does not correspond to the original layout of our algorithm, but lets us efficiently explore some limitations of this approach. We have learned that for interpreting the results of our approach, domain knowledge is advantageous as we have identified incomplete and inconsistent configuration interfaces as possible limiting factors. We acknowledge that, while our approach performs well as an automated pipeline in a controlled environment, for a real-world applications, a semi-automated setup with additional manual measurement prioritization and knowledge about configuration constraints would improve the applicability of our approach.} This is a promising further avenue of future work.

@@

	\section{Related Work}\label{sec:rw}
	\paragraph{Performance-relevant Commits} Identifying performance-changing commits is a prevalent challenge in regression testing, often applying change point detection algorithms on batches of performance observations~\cite{cityIdentifying2014,daly_industry_2020}. A way to reduce testing overhead is to employ only a fraction of performance observations or prioritize commits. Sandoval Alcocer et al. assess performance for uniformly selected commits to estimate the risk of a performance change introduced by unobserved commits~\cite{sandoval_alcocer_learning_2016,alcocer_prioritizing_2020}. Huang et al. estimate the risk of performance changes solely based on static analysis of individual commits to prioritize commits for regression testing without assessing performance~\cite{huang_performance_2014}. Mühlbauer et al. use Gaussian Processes (GP) and iterative sampling to learn performance histories with few performance observations~\cite{muhlbauer_accurate_2019}. They extract performance changing commits by applying change point detection algorithms to learned performance histories. {\color{red}Nonetheless, they do not associate change points with individual configuration options or interactions.} All of the above approaches reduce testing overhead, but do not address the performance across different configurations of software systems. Our approach incorporates performance changes across different configurations by leveraging similarities in related configuration's performance histories.
	\paragraph{Performance of Configurable Software Systems} Associating configuration options and interactions with performance a conceptual basis for our work. There exists extensive work on modeling configuration options as features for machine learning, such as classification and regression trees~\cite{guoVariabilityawarePerformancePrediction2013,sarkarCostEfficientSamplingPerformance,nairUsingBadLearners2017}, multi-variable regression~\cite{siegmundPerformanceinfluenceModelsHighly2015}, and deep neural networks~\cite{haDeepPerf2019}. In contrast to learning performance for arbitrary configurations, our approach aims at explaining performance change, and subsequently, a change in a configuration option's or interaction's influence on performance over time. A similar scenario to ours has been studied by Jamshidi et al.~\cite{jamishidi_transfer_2017,jamshidi_transfer_gp_2017,jamshidi_learning_2018}. Between software versions, the performance influence of only few configuration options changes. The authors propose two approaches for learning a transfer function between performance models of different environments, such as versions, based on GPs~\cite{jamshidi_transfer_gp_2017} and progressive shrinking the configuration space~\cite{jamshidi_learning_2018}. We see great potential for future work in applying GPs to our problem statement as they have been successfully applied for the two orthogonal perspectives of time and configuration space.
